Pipit: Reactive Systems in \fstar ~(Extended Abstract)

### 2023-05-19 2023pipit/Main.tex
tried to add more related work but it will not fit!

Pipit is an \emph{embedded} language, like Bedrock~\cite{chlipala2013bedrock}: in this example, the stream type denotes an actual Pipit expression, while the list type and its associated pattern match is part of the \fstar{} meta-language.
To get the real Pipit program, we need to apply the filter to a concrete list of coefficients:

@@

Recursive streams, which can refer to previous values of the stream itself, are defined using the fixpoint operator ($\xrec{x}{e[x]}$); the syntax $e[x]$ means that the variable $x$ can occur in $e$.
% To ensure that streams are productive,
As in Lustre, recursive streams can only refer to their previous values and must be \emph{guarded} by a delay: the stream $(\xrec{x}{\xfby{0}{(x + 1)}})$ is well-defined, but stream $(\xrec{x}{x + 1})$ is invalid and has no computational interpretation.
This form of recursion differs slightly from standard Lustre, which uses a set of mutually-recursive bindings.
We use this form to define a substitution-based operational semantics that is syntax-directed, as opposed to the mutually-recursive form in \citet{caspi1995functional} which is not syntax directed.
The syntax-directed semantics simplifies the proof of determinism; we believe it has simplified other necessary proofs too and will perform further evaluation.
Although we cannot express mutually-recursive bindings in the core syntax here, we can express them as a notation on the surface syntax at the expense of potentially duplicating expressions.

>>>>

Pipit is an \emph{embedded} language, like Bedrock~\cite{chlipala2013bedrock}: in this example, the stream type denotes an actual Pipit expression, while the list type and its associated pattern match are part of the \fstar{} meta-language.
To get the real Pipit program, we need to apply the filter to a concrete list of coefficients:

@@

Recursive streams, which can refer to previous values of the stream itself, are defined using the fixpoint operator ($\xrec{x}{e[x]}$); the syntax $e[x]$ means that the variable $x$ can occur in $e$.
% To ensure that streams are productive,
As in Lustre, recursive streams can only refer to their previous values and must be \emph{guarded} by a delay: the stream $(\xrec{x}{\xfby{0}{(x + 1)}})$ is well-defined, but stream $(\xrec{x}{x + 1})$ is invalid and has no computational interpretation.
This form of recursion differs slightly from standard Lustre, which uses a set of mutually-recursive bindings.
We use this form to define a substitution-based operational semantics that is syntax-directed, as opposed to the mutually-recursive form in \citet{caspi1995functional} which is not syntax-directed.
Our semantics has a simpler proof of determinism; we believe it has simplified other necessary proofs too and will perform further evaluation.
Although we cannot express mutually-recursive bindings in the core syntax here, we can express them as a notation on the surface syntax at the expense of potentially duplicating expressions.
